---
title: withr 2.1.0
date: '2017-11-15'
slug: withr-2.1.0
author: Jim Hester
categories: [package]
description: >
  withr 2.1.0 is now available on CRAN.
photo:
  url: https://unsplash.com/photos/V1YEEItoOTE
  author: Suganth
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
)
library(withr)
```

[withr 2.1.0](http://withr.r-lib.org) is now available on CRAN!
[withr](http://withr.r-lib.org) makes working with global state in R safer and
less error prone. It also has no non-base dependencies so is easy to include in
packages.

Install the latest version with:
```{r, eval = FALSE}
install.packages("withr")
```

# Why withr?

Many functions in R modify global state in some fashion. Some common examples
are `par()` for graphics parameters, `dir()` to change the current directory
and `options()` to set a global option. Using these global options is useful
when using R interactively, because you can set them early in your
experimentation and they will remain set for the duration of the session.
However this makes programming with these settings difficult, because in a
pure function you should always reset the previous state when the function
completes if you change a global setting.

One common idiom for dealing with this problem is to save the current state,
make your change, then reset the previous state.

```{r}
par("col" = "black")
my_plot <- function(new) {
  old <- par(col = "red", pch = 19)
  plot(mtcars$hp, mtcars$wt)
  par(old)
}
my_plot()
par("col")
```

However one issue when there is an error before you are able to reset the
options.

```{r, error = TRUE}
par("col" = "black")
my_plot <- function(new) {
  old <- par(col = "red", pch = 19)
  plot(mtcars$hpp, mtcars$wt)
  par(old)
}
my_plot()
par("col")
```

Using the base function `on.exit()` is the robust solution to this problem,
which will run the resetting code when the block is exited, regardless if it
exits normally or with an error.

```{r, error = TRUE}
par("col" = "black")
my_plot <- function(new) {
  old <- par(col = "red", pch = 19)
  on.exit(par(old))
  plot(mtcars$hpp, mtcars$wt)
}
my_plot()
par("col")
```

However this solution is somewhat cumbersome to work with. You
need to remember to use an `on.exit()` call after each stateful call. In
addition by default each `on.exit()` action will overwrite any previous
`on.exit()` action in the same block (using `add = TRUE` prevents this). It is also
not possible to have this cleanup performed before the block has finished.

[withr](http://withr.r-lib.org) is a solution to these issues. It defines a
[large set of
functions](http://withr.r-lib.org/#withr---run-code-with-modified-state) for
dealing with global settings in R, such as `with_par()`. These functions set one of
the global settings for the duration of a block of code, then automatically
reset it after the block is completed.

```{r}
par("col" = "black")
my_plot <- function(new) {
  with_par(list(col = "red", pch = 19),
    plot(mtcars$hp, mtcars$wt)
  )
  par("col")
}
my_plot()
par("col")
```

In addition to the `with_*` functions there are `local_*` variants which simply
last until the end of the block they are included in. These work similar to
`on.exit()`, but you can set the options in one call rather than two.

```{r}
par("col" = "black")
my_plot <- function(new) {
  local_par(list(col = "red", pch = 19))
  plot(mtcars$hp, mtcars$wt)
}
my_plot()
par("col")
```

# Graphics devices

There are now a comprehensive set of functions to deal with R's builtin
[graphics devices](http://withr.r-lib.org/reference/devices.html).

These functions open a new graphics device, run some code, then automatically
close the device. This is very useful when printing multiple plots to the same
device.

```{r}
with_pdf("test.pdf", width = 7, height = 5, {
  plot(mtcars$hp, mtcars$mpg)
  plot(mtcars$wt, mtcars$mpg)
})
```

![](test.pdf)

Thanks to [Richard Cotton's](https://github.com/richierocks) great [pull
request](https://github.com/r-lib/withr/pull/37) for this feature!

## Connections

There are two new functions for cleaning up connections in R.
`with_connection()` allows you to automatically close R's file connections.
Here we create a writable file connection, write some lines to it with
`with_connection()`, then open a read-only connection and read the file using
`local_connection()`.

```{r}
with_connection(list(con = file("temp", "w")), {
  writeLines(c("foo", "bar"), con)
})
read_temp <- function() {
  readLines(local_connection(file("temp", "r")))
}
read_temp()
```
```{r, echo = FALSE}
unlink("temp")
```

`with_db_connection()` provides [DBI](http://rstats-db.github.io/DBI)
connections to Relational Database Management Systems which automatically call
`DBI::dbDisconnect()`. Here we create a new [SQLite](https://www.sqlite.org/)
database, connect to it with `with_db_connection()`, and write a new table to
it. We then create another connection with `local_db_connection()` and read
from the table.

```{r}
db <- tempfile()
with_db_connection(
  list(con = DBI::dbConnect(RSQLite::SQLite(), db)), {
    DBI::dbWriteTable(con, "mtcars", mtcars)
})

head_db_table <- function(...) {
  con <- local_db_connection(DBI::dbConnect(RSQLite::SQLite(), db))
  head(DBI::dbReadTable(con, "mtcars"), ...)
}
head_db_table()
unlink(db)
```

## Packages / Namespaces / Environments

Three new functions allow you to temporary attach a package, namespace or
environment in a block of code.

```{r}
with_package("lattice", {
  xyplot(y ~ x, data.frame(x = -2:2, y = dnorm(-2:2)))
})
```

`with_namespace()` attaches the package namespace, so all objects (including
unexported ones) are available. Here `vlapply` is an unexported functions from withr.

```{r}
with_namespace("withr", {
  vlapply(1:5, function(x) x %% 2 == 0)
})
```

`with_environment()` attaches any environment, detaching it in the same fashion
as the other functions.


## Other changes

There are many other bug fixes and other minor improvements in this release.
You can see a complete list in the [release
notes](https://github.com/tidyverse/withr/releases/tag/v2.1.0).

A big thanks goes to all the community members who contributed code and opened
issues since the last release!

```{r, eval = FALSE, include = FALSE}
x <- gh::gh("/repos/:owner/:repo/issues", owner = "r-lib", repo = "withr", since = "2017-06-16", state = "all")
users <- unique(purrr::map_chr(x, c("user", "login")))
clipr::write_clip(glue::collapse(glue::glue("[\\@{users}](https://github.com/{users})"), ", ", last = ", and "))
```

[\@QuLogic](https://github.com/QuLogic),
[\@krlmlr](https://github.com/krlmlr),
[\@hadley](https://github.com/hadley),
[\@wlandau-lilly](https://github.com/wlandau-lilly),
[\@jimhester](https://github.com/jimhester),
[\@kevinushey](https://github.com/kevinushey),
and [\@richierocks](https://github.com/richierocks)
